Arhangelskii's \(\alpha\)-principles and selection games

### 2019-10-30 articles/selection-alpha-games/article.tex
add another citation
add narrative

\title{Selection games and Arhangelskii's convergence principles}

@@

\begin{abstract}
We prove the things.
\end{abstract}

@@

%
%\section{Introduction}
%
%\begin{definition}
%  The \term{selection game} \(\schStrongSelGame{\mc A}{\mc B}\) 
%  is an \(\omega\)-length game involving Players \(\plI\) and \(\plII\). 
%  During round \(n\), \(\plI\) chooses
%  \(A_n\in\mc A\), followed by \(\plII\) choosing \(B_n\in A_n\).
%  Player \(\plII\) wins in the case that \(\{B_n:n<\omega\}\in\mc B\),
%  and Player \(\plI\) wins otherwise.
%\end{definition}
%
%\begin{definition}
%  The \term{selection game} \(\schSelGame{\mc A}{\mc B}\) 
%  is an \(\omega\)-length game involving Players \(\plI\) and \(\plII\). 
%  During round \(n\), \(\plI\) chooses
%  \(A_n\in\mc A\), followed by \(\plII\) choosing \(B_n\in [A_n]^{<\aleph_0}\).
%  Player \(\plII\) wins in the case that \(\bigcup\{B_n:n<\omega\}\in\mc B\),
%  and Player \(\plI\) wins otherwise.
%\end{definition}
%
\section{Clontz results}

@@

\begin{definition}
Let \(\Gamma_X\) be the collection of open \term{\(\gamma\)-covers} \(\mc U\) of \(X\),
that is, infinite open covers of \(X\) such that for each \(x\in X\),
\(\{U\in\mc U:x\in U\}\) is cofinite in \(\mc U\).
\end{definition}

\begin{definition}
Let \(\Gamma_{X,x}\) be the collection of non-trivial sequences \(S\subseteq X\) converging to \(x\),
that is, infinite subsets of \(X\) such that for each neighborhood \(U\) of \(x\),
\(S\cap U\) is cofinite in \(S\).
\end{definition}

It follows that \(\Gamma_X,\Gamma_{X,x}\) are both \(\Gamma\)-like. We also require
the following.

@@

\begin{theorem}
Let \(\mc B\) be \(\Gamma\)-like. Then \(\alpha_1(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_{cf}(\mc A,\mc B)\).
\end{theorem}

@@

\begin{proof}
We first assume \(\alpha_2(\mc A,\mc B)\) and let \(A_n\in\mc A\) for \(n<\omega\)
define a predetermined strategy for \(\pl I\).
We may apply \(\alpha_2(\mc A,\mc B)\) to choose \(B\in\mc B\) such that
\(|A_n\cap B|\geq\aleph_0\). We may then choose \(a_n\in(A_n\cap B)\setminus\{a_i:i<n\}\)
for each \(n<\omega\). It follows that \(B'=\{a_n:n<\omega\}\in\mc B\) since
\(B'\) is an infinite subset of \(B\in\mc B\); therefore \(A_n\) does not define
a winning predetermined strategy for \(\plI\).

@@

@@

We further note that the above proof technique could be
used to establish that perfect-information and
Markov winning strategies for \(\plII\) in
\(G_{fin}(\mc A,\mc B)\) may be improved to
be valid in \(G_{<2}(\mc A,\mc B)\), provided
\(\mc B\) is \(\Gamma\)-like. As such,
\(G_{<2}(\mc A,\mc B)\) and \(G_{fin}(\mc A,\mc B)\)
are effectively equivalent games under this hypothesis.
%\begin{theorem}
%Let \(\mc B\) be \(\Gamma\)-like. Then \(\plII\win G_{<2}(\mc A,\mc B)\)
%if and only if \(\plII\win G_{fin}(\mc A,\mc B)\).
%\end{theorem}

%\begin{theorem}
%Let \(\mc B\) be \(\Gamma\)-like. Then \(\plII\markwin G_{<2}(\mc A,\mc B)\)
%if and only if \(\plII\markwin G_{fin}(\mc A,\mc B)\).
%\end{theorem}

\begin{theorem}\label{pawlikowskii}
Let \(\mc A\) be almost-\(\Gamma\)-like and \(\mc B\) be \(\Gamma\)-like.
Then \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\), and
\(\plI\win G_1(\mc A,\mc B)\) if and only if
\(\plI\prewin G_1(\mc A,\mc B)\).
\end{theorem}

@@

Note that the assumption in Theorem \ref{pawlikowskii} that \(\mc A\)
be almost-\(\Gamma\)-like cannot be omitted. In
[todo cite Clontz k-tactics in Gruenhage game]
an example of a space and point where \(\plI\win G_1(\mc A,\mc B)\) but
\(\plI\notprewin G_1(\mc A,\mc B)\) is given, where \(\mc A\) is
the set of open neighborhoods of the given point 
(which are all uncountable), 
and \(\mc B\) is the set of converging sequences to that point.
(Note that \(G_1(\mc A,\mc B)\) is called \(Gru_{O,P}(X,x)\) in that
paper. In fact, more is shown: \(\plI\) has a winning perfect-information
strategy, but any strategy that only uses 
the most recent \(k\) moves of \(\plII\) and the round number
can be defeated, where \(k\) is any natural number.)

@@

\begin{corollary}
Let \(\mc B\) be \(\Gamma\)-like and \(\mc B\subseteq\mc A\).
Then \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\), and
\(\plI\win G_{1}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{1}(\mc A,\mc B)\).
\end{corollary}
\begin{proof}
Assuming \(\plI\notprewin G_{fin}(\mc A,\mc B)\), we have
\(\plI\notwin G_{fin}(\mc A,\mc B)\) by Proposition \ref{auto-asl}
and Theorem \ref{pawlikowskii}.

@@

This corollary generalizes e.g. Theorems 26 and 30 of [cite Scheepers 1996 Ramsey]
and Theorem 5 of [cite MR2119791].

>>>>

\title{Arhangelskii's \(\alpha\)-principles and selection games}

@@

\begin{abstract}
Arhangelskii's convergence properties \(\alpha_2\) and \(\alpha_4\)
may be characterized in terms of Scheeper's selection games.
We generalize these folklore results to hold for more general collections.
\end{abstract}

@@

The following characterizations were given as Definition 1 by Kocinac in
[cite Kocinac selection principles related].

\begin{definition}
\term{Arhangelskii's \(\alpha\)-principles} \(\alpha_i(\mc A,\mc B)\) are defined as follows
for \(i\in\{1,2,3,4\}\).
Let \(A_n\in\mc A\) for all \(n<\omega\); then there exists \(B\in\mc B\) such that:
\begin{itemize}
\item[\(\alpha_1\):] \(A_n\cap B\) is cofinite in \(A_n\) for all \(n<\omega\).
\item[\(\alpha_2\):] \(A_n\cap B\) is infinite for all \(n<\omega\).
\item[\(\alpha_3\):] \(A_n\cap B\) is infinite for infinitely-many \(n<\omega\).
\item[\(\alpha_4\):] \(A_n\cap B\) is non-empty for infinitely-many \(n<\omega\).
\end{itemize}
\end{definition}

When \((\mc A,\mc B)\) is omitted, it is assumed that \(\mc A=\mc B\) is the
collection \(\Gamma_{X,x}\) of sequences converging to some point \(x\in X\),
as introduced by Arhangelskii in [cite Arhangelskii frequency spectrum].
Provided \(\mc A\) only contains infinite sets, it's easy to see that
\(\alpha_n(\mc A,\mc B)\) implies \(\alpha_{n+1}(\mc A,\mc B)\).

We aim to relate these to the following games.

\begin{definition}
  The \term{selection game} \(G_1(\mc A,\mc B)\) (resp. \(G_{fin}(\mc A,\mc B)\))
  is an \(\omega\)-length game involving Players \(\plI\) and \(\plII\). 
  During round \(n\), \(\plI\) chooses
  \(A_n\in\mc A\), followed by \(\plII\) choosing \(a_n\in A_n\)
  (resp. \(F_n\in[A_n]^{<\aleph_0}\)).
  Player \(\plII\) wins in the case that \(\{a_n:n<\omega\}\in\mc B\)
  (resp. \(\bigcup\{F_n:n<\omega\}\in\mc B\)),
  and Player \(\plI\) wins otherwise.
\end{definition}

Such games are well-represented in the literature; see [cite Scheepers combinatorics ramsey]
for example.
We will also consider the similarly-defined games \(G_{<2}(\mc A,\mc B)\)
(\(\plII\) chooses 0 or 1 points from each choice by \(\plI\)) and \(G_{cf}(\mc A,\mc B)\)
(\(\plII\) chooses cofinitely-many points).

\begin{definition}
  Let \(P\) be a player in a game \(G\). \(P\) has a \term{winning strategy}
  for \(G\), denoted \(P\win G\), if \(P\) has a strategy that defeats every
  possible counterplay by their opponent. If a strategy only relies on the
  round number and ignores the moves of the opponent, the strategy is said
  to be \term{predetermined}; the existence of a predetermined winning strategy
  is denoted \(P\prewin G\).
\end{definition}

We briefly note that the statement \(\plI\notprewin G_\star(\mc A,\mc B)\)
is often denoted as the \term{selection principle} \(S_\star(\mc A,\mc B)\).

The equivalence of \(\alpha_2(\Gamma_{X,x}\Gamma_{X,x})\) and 
\(\plI\notprewin G_1(\Gamma_{X,x},\Gamma_{X,x})\) was briefly asserted by Sakai
in the introduction of [cite Sakai sequence selection properties]; the similar
equivalence of \(\alpha_4(\Gamma_{X,x}\Gamma_{X,x})\) and 
\(\plI\notprewin G_{fin}(\Gamma_{X,x},\Gamma_{X,x})\) seems to be folklore.
In fact, these relationships hold in more generality.

\section{Main Results}

\begin{definition}
Let \(\Gamma_{X,x}\) be the collection of non-trivial sequences \(S\subseteq X\) converging to \(x\),
that is, infinite subsets of \(X\setminus\{x\}\) such that for each neighborhood \(U\) of \(x\),
\(S\cap U\) is cofinite in \(S\).
\end{definition}

\begin{definition}
Let \(\Gamma_X\) be the collection of open \term{\(\gamma\)-covers} \(\mc U\) of \(X\),
that is, infinite open covers of \(X\) such that \(X\not\in \mc U\) and for each \(x\in X\),
\(\{U\in\mc U:x\in U\}\) is cofinite in \(\mc U\).
\end{definition}

The similarity in nomenclature follows from the observation that every non-trivial sequence in
\(C_p(X)\) converging to the zero function \(\vec{0}\) naturally defines a corresponding
\(\gamma\)-cover in \(X\), see e.g. Theorem 4 of
[Scheepers a sequential property and covering property].

Note that by these definitions, convergent sequences (resp. \(\gamma\)-covers) may be uncountable,
but any infinite subset of either would remain a convergent sequence (resp. \(\gamma\)-cover),
in particular, countably infinite subsets. We capture this idea as follows.

@@

We also require the following.

@@

We are now able to prove a few general equivalences between \(\alpha\)-princples
and selection games.

@@

\begin{proof}
We first assume \(\alpha_2(\mc A,\mc B)\) and let \(A_n\in\mc A\) for \(n<\omega\)
define a predetermined strategy for \(\plI\).
We may apply \(\alpha_2(\mc A,\mc B)\) to choose \(B\in\mc B\) such that
\(|A_n\cap B|\geq\aleph_0\). We may then choose \(a_n\in(A_n\cap B)\setminus\{a_i:i<n\}\)
for each \(n<\omega\). It follows that \(B'=\{a_n:n<\omega\}\in\mc B\) since
\(B'\) is an infinite subset of \(B\in\mc B\); therefore \(A_n\) does not define
a winning predetermined strategy for \(\plI\).

@@

This shows that \(\plII\) gains no advantage from picking more than one
point per round. This in fact only depends on \(\mc B\) being
\(\Gamma\)-like, which we formalize in the following results.

@@

We note that the above proof technique could be
used to establish that perfect-information and
limited-information strategies for \(\plII\) in
\(G_{fin}(\mc A,\mc B)\) may be improved to
be valid in \(G_{<2}(\mc A,\mc B)\), provided
\(\mc B\) is \(\Gamma\)-like. As such,
\(G_{<2}(\mc A,\mc B)\) and \(G_{fin}(\mc A,\mc B)\)
are effectively equivalent games under this hypothesis.

\begin{theorem}\label{pawlikowskii}
Let \(\mc A\) be almost-\(\Gamma\)-like and \(\mc B\) be \(\Gamma\)-like. Then
\begin{itemize}
\item \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\), and
\item \(\plI\win G_1(\mc A,\mc B)\) if and only if
\(\plI\prewin G_1(\mc A,\mc B)\).
\end{itemize}
\end{theorem}

@@

Note that the assumption in Theorem \ref{pawlikowskii} that \(\mc A\)
be almost-\(\Gamma\)-like cannot be omitted. In
[todo cite Clontz k-tactics in Gruenhage game]
an example of a space \(X^*\) and point \(\infty\in X^*\) 
where \(\plI\win G_1(\mc A,\mc B)\) but
\(\plI\notprewin G_1(\mc A,\mc B)\) is given, where \(\mc A\) is
the set of open neighborhoods of \(\infty\) 
(which are all uncountable), 
and \(\mc B\) is the set \(\Gamma_{X^*,\infty}\) of sequences converging to that point.
(Note that \(G_1(\mc A,\mc B)\) is called \(Gru_{O,P}(X^*,\infty)\) in that
paper, and an equivalent game \(Gru_{K,P}(X)\) is what is directly
studied. In fact, more is shown: \(\plI\) has a winning perfect-information
strategy, but for any natural number \(k\), any strategy that only uses 
the most recent \(k\) moves of \(\plII\) and the round number
can be defeated.)

While \(\mc A\) is often not almost-\(\Gamma\)-like in general,
it may satisfy that property in combination with the selection principles
being considered.

@@

Note that in the previous result, 
\(\plI\notprewin G_{fin}(\mc A,\mc B)\) could be weakened
to the choice principle \({\mc A}\choose{\mc B}\): for every
member of \(\mc A\), there is some countable subset belonging to \(\mc B\).

\begin{corollary}
Let \(\mc B\) be \(\Gamma\)-like and \(\mc B\subseteq\mc A\).
Then 
\begin{itemize}
\item \(\plI\win G_{fin}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{fin}(\mc A,\mc B)\), and
\item \(\plI\win G_{1}(\mc A,\mc B)\) if and only if
\(\plI\prewin G_{1}(\mc A,\mc B)\).
\end{itemize}
\end{corollary}
\begin{proof}
Assuming \(\plI\notprewin G_{fin}(\mc A,\mc B)\), we have
\(\plI\notwin G_{fin}(\mc A,\mc B)\) by Proposition \ref{auto-asl}
and Theorem \ref{pawlikowskii}.

@@

This corollary generalizes e.g. Theorems 26 and 30 of [cite Scheepers 1996 Ramsey],
Theorem 5 of [cite MR2119791], and Corollary 36 of [cite Clontz dual games].

In summary, using the selection principle notation \(S_\star(\mc A,\mc B)\):

\begin{corollary}
Let \(\mc B\) be \(\Gamma\)-like and \(\mc B\subseteq\mc A\).
Then 
\begin{itemize}
\item \(\plI\notwin G_{fin}(\mc A,\mc B)\) if and only if
\(S_{fin}(\mc A,\mc B)\) if and only if
\(\alpha_2(\mc A,\mc B)\), and
\item \(\plI\notwin G_{1}(\mc A,\mc B)\) if and only if
\(S_{1}(\mc A,\mc B)\) if and only if
\(\alpha_4(\mc A,\mc B)\).
\end{itemize}
\end{corollary}

\section{Conclusion}

We conclude with the following easy result, and a couple questions.

\begin{proposition}
Let \(\mc B\) be \(\Gamma\)-like. Then \(\alpha_1(\mc A,\mc B)\) holds if and only
if \(\plI\notprewin G_{cf}(\mc A,\mc B)\).
\end{proposition}

\begin{question}
Is there a game-theoretic characterization of \(\alpha_3(\mc A,\mc B)\)?
\end{question}

Noting that \(\plI\win G_1(\Gamma_X,\Gamma_X)\) if and only if
\(\plI\win G_{fin}(\Gamma_X,\Gamma_X)\) [cite Kocinac], but
the same is not true of \(G_\star(\Gamma_{X,x},\Gamma_{X,x})\)
(i.e. there are \(\alpha_4\) spaces that are not \(\alpha_2\)
[cite Arhangelskii]),
we also ask the following.

\begin{question}
Is there an elegant condition on \(\mc A,\mc B\) guaranteeing
\(\plI\win G_1(\mc A,\mc B)\Rightarrow\plI\win G_{fin}(\mc A,\mc B)\)?
\end{question}
